---
output:
  html_document: default
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.pos = "!H", out.extra = "")
library("GillespieSSA")
library("ggplot2")
library("patchwork")
library("ggpmisc")
library("stringr")
# SOURCE .RDATA FILE of Lias matrices (.csv)
load(file = "lia_matrices.RData")
source(file = "C:/Users/sedar/Documents/Seda_Radoykova/UCL/1_BIOL0044 Th_Yr_Pr/stochastic_simulations/dependencies/plt_fnc.R") # plotting function
source(file = "C:/Users/sedar/Documents/Seda_Radoykova/UCL/1_BIOL0044 Th_Yr_Pr/stochastic_simulations/dependencies/react_prod_names.R")
#load("co2_fix_mat.RData")
```


### Reactant/product names (i.e. reactions)

```{r}
# getting reactant and product names 
reac_prod = read.table(file = "clipboard", header = F, sep = "\t")
react = react_prod_names(reac_prod$V1)
prod = react_prod_names(reac_prod$V2)
all_spec = unique(unlist(c(react, prod)))
#save(react, prod, all_spec, st_ch_mat, st_ch_df, file = "co2_fix_mat.RData")
```

### State change matrix

```{r}
st_ch_mat = matrix(
  data = rep(0, (length(all_spec) * length(react))),
  nrow = length(all_spec),
  ncol = length(react)
)

for (i in seq_along(react)) {
  for (j in seq_along(react[[i]])) {
    tmp_row = which(react[[i]][j] == all_spec)
    st_ch_mat[tmp_row,i] = -1 
  }
  for (j in seq_along(prod[[i]])) {
    tmp_row = which(prod[[i]][j] == all_spec)
    st_ch_mat[tmp_row,i] = +1 
  }
}

st_ch_mat
st_ch_df = as.data.frame(x = st_ch_mat, row.names = all_spec)
#colnames(st_ch_df) = tolower(str_replace_all(some_rxn_names, " ", "_"))
#print.data.frame(st_ch_df)
#write.csv(st_ch_df, "co2_fix.csv")
```


### Reaction formulae

```{r}
# reactant*reactant 
ks = vector(mode = "character", length = length(react))
a = 0
for (i in seq_along(react)) {
  a = a + 1
  ks[a] = str_c(react[[i]], collapse = "*")
}


#k_1*reactant*reactant
full_ks = vector(mode = "character", length = length(react))
#k_whatever
k_char = vector("character", length(ks))
for (i in seq_along(ks)) {
  #cat("\"k_", i, "*", ks[i], "\",\n", sep = "")
  full_ks[i] = str_c("k_", i, "*", ks[i], collapse = "")
  k_char[i] = str_c("k_", i, collapse = "")
}
```

```{r}
k_char
ks
full_ks
```


### Run model here  

```{r}
params_k = c(10, 19.9, 23.5, 0.061, 0.869)
names(params_k) = k_char
```

```{r}
#initial concentrations of reactants and products involved 
# named vector of quantities 
nmol = 5000
x_co2_fix = c(rep(nmol, 2), rep(0,6))
names(x_co2_fix) = all_spec
```

```{r}
# reactions rates / probability of reaction
a_co2_fix = full_ks
```

```{r}
# state change matrix
mat_co2_fix = st_ch_mat

# simulation
out_co2_fix <- ssa(x0 = x_co2_fix, a = a_co2_fix, 
                      nu = mat_co2_fix, parms = params_k, 
                      tf = 0.0005, method = ssa.d())
```

```{r}
# plot output
co2_fix = plt_fnc(
  to_plot = colnames(out_co2_fix[["data"]])[-1], 
  data = out_co2_fix[["data"]], col = "red"
  )
```


### Lia's simulations

```{r}
parms_k <- c(
  k_1 = 10,
  k_2 = 19.9,
  k_3 = 23.5,
  k_4 = 0.869,
  k_5 = 0.061
) 
nmol <- 5000

#initial concentrations of reactants and products involved 
x_c_fix <- c(CO2 = nmol, H2 = nmol, formic_acid = 0,
        formaldehyde = 0, methanol = 0, CO = 0, acetic_acid = 0, H2O = 0) 

a_c_fix <- #carbon fixation
  c(
    "k_1*CO2*H2",
    "k_2*formic_acid*H2",
    "k_3*formaldehyde*H2",
    "k_4*methanol*CO",
    "k_5*CO2*H2"
  )

# state change matrix

mat_c_fix <- as.matrix(csvs_matrices$no_cofactors[,2:6])

out_c_fix <- ssa(x0 = x_c_fix, a = a_c_fix, 
                      nu = mat_c_fix, parms = parms_k, 
                      tf = 0.0005, method = ssa.d())
```

``` {r carbon fixation new, warning=FALSE, echo=FALSE}
parms_k
params_k

#initial concentrations of reactants and products involved 
x_co2_fix
x_c_fix 

a_c_fix 
a_co2_fix

# state change matrix

mat_c_fix 
mat_co2_fix
```

```{r}
fwd = plt_fnc(
  to_plot = colnames(out_c_fix[["data"]])[-1], 
  data = out_c_fix[["data"]], col = "black"
  )
#for (i in 1:16) cat("p1[[", i, "]] + ", sep = "")
# fwd[[1]] + fwd[[2]] + fwd[[3]] + fwd[[4]] + fwd[[5]] + fwd[[6]] + fwd[[7]] + fwd[[8]]
```


# Manually implemented simualtions - RED
```{r fig.height=3.5}
table = data.frame(reaction = full_ks, parameters = str_c(k_char, " = ", params_k))

fig = co2_fix[[1]] + co2_fix[[2]] + co2_fix[[3]] + co2_fix[[4]] + co2_fix[[5]] + co2_fix[[6]] + co2_fix[[7]] + co2_fix[[8]]

ggsave(fig + annotate(geom = "table", x = 5, y = 2, label = list(table), 
                      vjust = 0.3, hjust = -0.11),
       file = paste0("seda_manual_sim.png"), width = 18, height = 13, units = "cm")

fig
```

# Lia's sim - BLACK

```{r fig.height=3.5}
fwd[[1]] + fwd[[2]] + fwd[[3]] + fwd[[4]] + fwd[[5]] + fwd[[6]] + fwd[[7]] + fwd[[8]]
```

PAY ATTENTION TO REACTION FORMULAE AND CORRESPONDING k-values!! 

- today's achievement: 
  - sort/organise code
  - compare to lia's code (all good, simulations match)
- to do: 
  - clean up code and make more independent R scripts to handle Stuart's input data
  - clean up simulation docs 


