---
output:
  html_document: default
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      warning = FALSE, out.extra = "")
library("GillespieSSA")
library("ggplot2")
library("patchwork")
library("ggpmisc")
library("stringr")
library("pse")
library("purrr")
library("cowplot")
library("rbenchmark")
source(file = "../dependencies/plt_fnc.R") # plotting function
```

This file contains an example of implementing a Latin-hypercube sampling based sensitivity analysis (SA) using the `pse` package framework. It starts out by spelling the reversible $CO_2$ fixation SSA model; generates a small LH sample to demonstrate the available functions. Then a more proper SA analysis is conducted (written up in the BIOL0044 project). Finally, a LH SA using the default settings is saved, as it generated interesting results. 


# Reversible CO2 fixation with equal forward and backward rates 

Writing the model out helps appreciate I/O relationships between `ssa()` and `LHS()`. 

```{r}
# initial concentrations of reactants and products involved 
nmol <- 5000

x_rev_c_fix <- c(
    CO2 = nmol,
    H2 = nmol,
    formic_acid = 0,
    formaldehyde = 0,
    methanol = 0,
    CO = 0,
    acetic_acid = 0,
    H2O = 0) 

a_rev_c_fix <- c( # reversible carbon fixation
                  "k_1*CO2*H2",
                  "k_1b*formic_acid",
                  "k_2*formic_acid*H2",
                  "k_2b*formaldehyde*H2O",
                  "k_3*formaldehyde*H2",
                  "k_3b*methanol",
                  "k_4*methanol*CO",
                  "k_4b*acetic_acid",
                  "k_5*CO2*H2",
                  "k_5b*CO*H2O")

# state change matrix
mat_rev_c_fix <- as.matrix(read.csv("../rev_c_fix/data/reversible_c_fix.csv", header = F)[-1])
```

```{r}
# parameters
parms_k <- c(k_1 = 10, k_1b = 10, 
             k_2 = 19.9, k_2b = 19.9, 
             k_3 = 23.5, k_3b = 23.5, 
             k_4 = 0.869, k_4b = 0.869, 
             k_5 = 0.061, k_5b = 0.061)

#output
out_rev_c_fix <- ssa(x0 = x_rev_c_fix, 
                      a = a_rev_c_fix, 
                      nu = mat_rev_c_fix,
                      parms = parms_k, 
                      tf = 0.0005, verbose = F, ignoreNegativeState = F)
#cat("Output dim =", dim(out_rev_c_fix$data))

rev_negl = plt_fnc(
  to_plot = colnames(out_rev_c_fix[["data"]])[-1], 
  data = out_rev_c_fix[["data"]], col = "darkred"
  )
#for (i in 1:16) cat("p1[[", i, "]] + ", sep = "")
fig = rev_negl[[1]] + rev_negl[[2]] + rev_negl[[3]] + rev_negl[[4]] + rev_negl[[5]] + rev_negl[[6]] + rev_negl[[7]] + rev_negl[[8]]
```

```{r}
# visualise results to check that sim is working 
fig
```



# PSE example with SSA model 

```{r}
# input parameters 
# params to be explored
factors = names(parms_k) # given as a vector
# their probability density distributions 
q = rep("qunif", 10) # given as a list 
# the arguments to be supplied to the probability density distributions
# given as a list of lists
q.arg = replicate(10, list(min = 1e-1, max = 1e2), simplify = F)
# names of results, to be used for plotting 
res.names = colnames(out_rev_c_fix$data)[-1]


# model 
# this function defines one iteration/run of the model 
# it takes SEPARTE vectors as arguments
# this is because the LH output is a matrix with cols = params and rows = vals 
### the arguments of this function would be modified to customise param exploration
rev_mod_run = function(k_1, k_1b, k_2, k_2b, k_3, 
                       k_3b, k_4, k_4b, k_5, k_5b){
  prms = c(k_1, k_1b, k_2, k_2b, k_3, k_3b, k_4, k_4b, k_5, k_5b)
  # since the GillespieR package takes in a named vector, make this here
  names(prms) = names(parms_k)
  # make sure the names of the parameters match 
  # with the input of the model run at the top of the document!
  run = ssa(
      x0 = x_rev_c_fix,
      a = a_rev_c_fix,
      nu = mat_rev_c_fix,
      parms = prms, # parms_k 
      # GillespieSSA package takes in a named vector of parameters 
      tf = 0.0005,
      ignoreNegativeState = F
    )
  # return the final state of the system (minus time data)
  return(as.array(run[["data"]][dim(run$data)[1],-1]))
  # can otherwise get individual columns out or some combination 
  ### editing what is returned could help tailor the sensitivity 
  ### analysis to a particular question of interest 
}


# wrapper function to run model 
# the wrapper should run the model given a data frame 
# with cols = params and rows = instances/values 
rev_mod = function(my_d) {
  names(my_d) = names(parms_k) 
  # NB name parmeters correctly, otherwise the model [ssa()] will not run 
  return(mapply(rev_mod_run, my_d[,1], my_d[,2], my_d[,3], 
                my_d[,4], my_d[,5], my_d[,6], 
                my_d[,7], my_d[,8], my_d[,9], my_d[,10]))
  }

# latin hypercube generation 
myLHS <- LHS(
  model = rev_mod, # model
  factors = factors, # parameters explored
  N = 50, # size of hypercube = number of samples generated 
  q = q, # quantile functions 
  q.arg = q.arg, # arguments for quantile functions
  nboot = 10, # bootstrap replicates; used for PRCC calculations
  res.names = res.names) # names used for visualisations
  #, repetitions = 5) # number of repetitions
```

```{r}
# checking that output and res.named match 
# as.array(out_rev_c_fix$data[dim(out_rev_c_fix$data)[1],-1])
# res.names
```

```{r}
### get data out for eyeballing 
# get.data(myLHS)
# get.repetitions(myLHS)
# get.results(myLHS)
```


# Visualise example

## ECDF 

```{r}
# examine the distribution of the model results
png("attempt2/ecdf1.png")
plotecdf(myLHS, stack = T)#, xlim=c(0,100))
  # stack = F gives several panels
    # applicable to multiple response vars
dev.off()
```

## Scatterplot

```{r}
# scatterplot of the result as a function of each parameter
pdf("attempt2/scatter.pdf")
plotscatter(myLHS)
  # index.res = which columns (aka factors) to be plotted
dev.off()
```

## PRCC 

PRCC is not very appropriate for stochastic models; PIC would be more appropriate. That being said, PIC is not easily visualisable the way PRCC is... 

```{r}
# partial rank correlation coefficient 
# assess the strength of the linear associations 
# between the result and each input parameter 
# after removing the linear effects of the other parameters
### used for deterministic models! 
purrr::map(1:7, function(x) {
  png(paste0("attempt1/prcc", x, ".png"))
  plotprcc(myLHS, index.res = x)
  dev.off()})
```

## PIC

Not easy to visualise. 

```{r}
# partial inclination coefficient
# stochastic equivalent of PRCC
# assess the sensitivity of the results to each parameter 

pic_an = pic(myLHS, nboot = 40)
# plt = myLHS
# plt$prcc = plts
# plotprcc(plt)
```

```{r}
colnames(myLHS$prcc[[1]]$PRCC)
plts = lapply(pic_an, function(x) x$prcc = x$pic)
```




# Proper PSE - included in project writeup

In the following sensitivity analysis, a Latin hypercube was used to vary parameters and examine the model outputs. 

All the parameters were set to be uniformly distributed, with range $[0.01,100]$. 

```{r}
load("LHS_proper/LHS_proper.RData")
```

## Check LH

This code helps check if the Latin hypercube samples behave as expected. 

```{r}
factors = names(parms_k) # given as a vector
q = rep("qunif", 10) # given as a list 
q.arg = replicate(10, list(min = 1e-2, max = 100), simplify = F)
res.names = colnames(out_rev_c_fix$data)[-1]
```


## LHS proper SA

```{r}
# benchmark PSE process time 
start_time <- Sys.time()
LHS_proper = LHS(model = rev_mod, factors = factors, 
                 N = 50, q = q, q.arg = q.arg, 
                 res.names = res.names, nboot = 50, repetitions = 10)
end_time <- Sys.time()
end_time - start_time
# Time difference of 14.81232 mins
```

```{r}
save(LHS_proper, file = "lhs2.RData")
```


## Plot parameters generated

```{r}
ggplot(reshape2::melt(LHS_proper$data), aes(value, fill = variable)) +
    geom_bar() +
    facet_wrap( ~ variable)
```

```{r}
# check data
#get.data(LHS_proper)

# save data 
# save(LHS_proper, file = "attempt2/LHS_proper.RData")
```

# Plots used in paper 

```{r}
# coloring ecdf plots - also using ggplot instead 
plotecdf(LHS_proper, stack = F, xlab = res.names[1], index.res = 1, lty = 2, col = "blue")

dat = as.vector(get.results(LHS_proper)[, 1])
g <- rep(1, each = dim(LHS_proper$res)[1])
Ecdf(dat, group = g, col = cols[1], xlab = res.names[1])
```

```{r}
cols = rcartocolor::carto_pal(8, "Safe")
# ggplot version 
dat = data.frame(get.results(LHS_proper))
colnames(dat) = res.names %>% stringr::str_replace(., "_", " ")
res_df = reshape2::melt(dat)
ecdfs_cols = ggplot(res_df, aes(value, color = factor(variable))) +
    stat_ecdf(geom = "line", size = 1.25) +
    facet_wrap(~ variable, ncol = 4, scales = "free_x") +
    scale_colour_manual(values = cols) +
    theme_cowplot() +
    labs(y = "Proportion <= x / [no unit]", x = "Number of molecules / [no unit]") +
    guides(color = guide_none()) 
```


# Plots - general

```{r}
# fix xlab ka = kb 
LHS_proper$factors = names(parms_k)
#LHS_proper$prcc %>% lapply(X = ., FUN = function(x) rownames(x[[1]]) = names(parms_k))
```


```{r}
# prcc 
png("LHS_proper/proper_prcc.png", res = 144, width = 1200, height = 800)
plotprcc(LHS_proper, ylab = res.names) # color every reaction possible 
dev.off()
```

```{r}
#LHS_proper$prcc[[8]]
molecule = rep(res.names, each = 10) %>% stringr::str_replace(., "_", " ")
all_data = lapply(LHS_proper$prcc, function(x) x$PRCC) %>% reduce(.f = rbind)
all_data$param = factor(row.names(all_data)[1:10])
all_data$molecule = factor(molecule, levels = unique(molecule))
rownames(all_data) = NULL
all_data = as.data.frame(all_data)
colnames(all_data)[4:5] = c("min_ci", "max_ci")
labs_k = sapply(all_data$param[seq(1,10,2)], function(x) c(paste(x), "")) %>% as.vector 
```

```{r}
png("LHS_proper/proper_prcc_GG.png", res = 144, width = 1200, height = 800)
ggplot(all_data, aes(x = as.factor(param), y = original,
                     ymin = min_ci, ymax = max_ci, color = molecule)) + 
    geom_pointrange(shape = 15) + 
    geom_hline(yintercept = 0, lty = 2) + 
    facet_wrap(~factor(molecule), nrow = 2) + 
    theme_cowplot() +
    scale_colour_manual(values = cols) +
    labs(y = "PRCC", x = "Reaction parameter") +
    guides(color = guide_none()) + 
    theme(axis.text.x = element_text(size = 10)) + 
    scale_x_discrete(labels = labs_k)
dev.off()
```


```{r}
sapply(all_data, typeof)  
```


```{r}
plotprcc(LHS_proper, ylab = res.names, col = cols[1], index.res = 1, main = "")
```


```{r}
# scatterplot of the result as a function of each parameter
p = 0
for (i in seq(1,10,by = 2)) {
  p = p + 1
  png(paste0("LHS_proper/scatter",p,".png"),
  plotscatter(LHS_proper, index.data = c(i, i + 1)))
  dev.off()
}
```

```{r}
# scatterplot of the result as a function of each parameter
plotss_scatter = map(seq(1,10,by = 2), function(x)
  return(eval(parse(
    text = paste0(
      "~{plotscatter(LHS_proper, index.data = c(", 
      paste(x), 
      ", ", paste(x + 1),"))}"
    )))))
```

```{r}
# scatter plots
png("LHS_proper/proper_scatters.png", res = 100, width = 500, height = 2200)
plot_grid(
  ggdraw(plotss_scatter[[1]]),
  ggdraw(plotss_scatter[[2]]),
  ggdraw(plotss_scatter[[3]]),
  ggdraw(plotss_scatter[[4]]),
  ggdraw(plotss_scatter[[5]]),
  nrow = 5, ncol = 1)
dev.off()
```

```{r}
ggsave("LHS_proper/ecdfs_cols.png", ecdfs_cols, 
       dpi = 100, units = "px", 
       width = 1000, height = 600)
```

```{r}
## ggplot scattter 
plotscatter(LHS_proper, index.data = c(1,2), )

# dat = data.frame(get.results(LHS_proper))
# colnames(dat) = res.names %>% stringr::str_replace(., "_", " ")
# res_df = reshape2::melt(dat)
params_lhs = data.frame(get.data(LHS_proper))
prm_lhs_df = reshape2::melt(params_lhs)
colnames(prm_lhs_df) = c("parameter", "param_val")
colnames(res_df) = c("molecule", "no_mol")

dat
res_df
params_lhs
prm_lhs_df


all_lms = data.frame(cbind(dat, params_lhs))
all_lms = reshape2::melt(all_lms, id.vars = colnames(params_lhs))
colnames(all_lms) = c(colnames(params_lhs), "molecule", "no_mol")
all_lms = reshape2::melt(all_lms, id.vars = c("molecule", "no_mol"))
colnames(all_lms) = c("molecule", "no_mol", "parameter", "param_val")

all_lms$molecule = stringr::str_replace(all_lms$molecule, "\\.", " ")
all_lms$molecule = factor(all_lms$molecule, levels = unique(all_lms$molecule))
```


```{r}
png("LHS_proper/proper_scatters_GG.png", res = 95, width = 1200, height = 1400)
ggplot(all_lms, aes(x = param_val, y = no_mol, color = molecule)) +
    geom_point(size = 0.5) +
    geom_smooth(method = 'lm', se = F) +
#    facet_grid(rows = vars(molecule), cols = vars(parameter), scales = "free", switch = "y") + 
#    facet_grid(rows = vars(parameter), cols = vars(molecule), scales = "free_y") + 
    facet_wrap(parameter ~ molecule, scales = "free_y", ncol = 8) + 
#    facet_wrap(molecule ~ parameter, scales = "free_y", ncol = 5) + 
    theme(strip.text = element_text(size = 8), 
          strip.text.y = element_text(size = 8),
          strip.text.x = element_text(size = 8)) + 
    theme_cowplot() +
    scale_colour_manual(values = cols) + 
    labs(x = "Parameter value / [no units]", y = "Number of molecules / [no units]") + 
    guides(color = guide_none()) 
dev.off()
```

```{r}
df = data.frame(molecule = dat[,1], k1 = params_lhs[,1])

ggplot(df, aes(x = k1, y = molecule)) +
  geom_point(size = 0.75) +
  geom_smooth(method = 'lm', se = F)
```

```{r}
# examine the distribution of the model results
plotss = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotecdf(LHS_proper, stack = F, xlab = res.names[",
      paste(x),
      "], index.res = ",
      paste(x),
      ")}"
    )
  ))))

# for (i in 1:8) {
#   cat(paste0("ggdraw(plotss[[",i,"]]), "))
#   ggdraw(plotss[[i]])
# }
```

```{r}
# ecdfs 
png("LHS_proper/proper_ecdfs.png", res = 100, width = 1000, height = 600)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),
  ggdraw(plotss[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]), 
  nrow = 2, ncol = 4)
dev.off()
```


```{r}
# example of getting globsl cv (coefficient of)
r <- get.results(LHS_proper)
sd(r) / mean(r) # global CV
plotcv(LHS_proper, stack = F, index.res = 1)
```

```{r}
plotss_cv = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotcv(LHS_proper, stack = F, index.res = ", paste(x),")}"
    )
  ))))
```


```{r}
# cv source code 
plotcv()

cv <- function(x, ...) {
    if (mean(x, ...) == 0) return(0)
    else return(sd(x, ...) / mean(x, ...))
}

pointw_cv = abs(apply(get.results(LHS_proper, FALSE), c(1, 2), cv))
global_cv = abs(apply(get.results(LHS_proper, TRUE), 2, cv))
quant = 0.99

for (i in 1) {
    m <- max(quantile(pointw_cv[, i], quant), 1.05 * global_cv)
    mi <- min(pointw_cv[, i], global_cv)
    Ecdf(pointw_cv[, i], xlim = c(mi, m), xlab = "pointwise cv")
    abline(v = global_cv[i], lwd = 2, lty = 3)
    if (m > 0.8 * max(pointw_cv[, i])) {pos = 2}
    else {pos = 4}
    text(x = global_cv[i], y = 0.1, label = "global cv", pos = pos
    )
}
```

```{r}
point_cv = as.data.frame(pointw_cv)
colnames(point_cv) = as.character(unique(all_lms$molecule))
point_cv_melt = reshape2::melt(point_cv)
colnames(point_cv_melt) = c("molecule", "point_cv")
point_cv_melt$molecule = factor(point_cv_melt$molecule, levels = colnames(point_cv))
global_cv_melt = data.frame(molecule = colnames(point_cv), global_cv = global_cv)
global_cv_melt$molecule = factor(global_cv_melt$molecule, levels = colnames(point_cv))
rownames(global_cv_melt) = NULL

point_cv_melt
global_cv_melt
```


```{r}
cols = rcartocolor::carto_pal(8, "Safe")
cv_cols = ggplot(point_cv_melt, aes(point_cv, color = factor(molecule))) +
    stat_ecdf(geom = "line", size = 1.25) +
    geom_vline(aes(xintercept = global_cv), global_cv_melt, lty = 2) +
    geom_text(aes(x = 0.9*global_cv, y = 0.5, label = "global cv"), data = global_cv_melt,
              text = element_text(size = 11, hjust = (0.5)), 
              color = "black", angle = 90) +
    facet_wrap(~molecule, ncol = 4, scales = "free_x") +
    scale_colour_manual(values = cols) +
    theme_cowplot() +
    guides(color = guide_none()) +
    labs(y = "Proportion <= x / [no unit]", x = "Pointwise CV") 
```

```{r}
ggsave("LHS_proper/cv_cols.png", cv_cols, 
       dpi = 100, units = "px", 
       width = 1000, height = 600)
```

```{r}
# both ecdf and pointwise cv
library("cowplot")
png("LHS_proper/ecdf_cv_col.png", res = 130, width = 1200, height = 1200)
plot_grid(ecdfs_cols, cv_cols, nrow = 2, ncol = 1, labels = "auto")
dev.off()
```



```{r}
library("cowplot")
png("LHS_proper/proper_cv.png", res = 100, width = 1000, height = 600)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),
  ggdraw(plotss[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]), 
  nrow = 2, ncol = 4, labels = res.names, vjust = 3.5)
dev.off()
```


```{r}
# mixed plot with ecdfs and cvs one below the other 
plotss = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotecdf(LHS_proper, stack = F, xlab = \"ecdf\", index.res = ",
      paste(x), ")}"
    )
  ))))

plotss_cv = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotcv(LHS_proper, stack = F, index.res = ", paste(x),")}"
    )
  ))))

res.names.new = c(rep(res.names[1:4], 2), rep(res.names[5:8], 2))

png("LHS_proper/proper_ecdf_cv.png", res = 100, width = 1000, height = 1200)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),  
  ggdraw(plotss[[4]]),
  ggdraw(plotss_cv[[1]]),
  ggdraw(plotss_cv[[2]]),
  ggdraw(plotss_cv[[3]]),
  ggdraw(plotss_cv[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]),
  ggdraw(plotss_cv[[5]]),
  ggdraw(plotss_cv[[6]]),
  ggdraw(plotss_cv[[7]]),
  ggdraw(plotss_cv[[8]]),
  nrow = 4, ncol = 4, labels = res.names.new, vjust = 3.5)
dev.off()
```

```{r}
# cv 
png("LHS_proper/proper_cv.png", res = 144, width = 1000, height = 600)
plotcv(LHS_proper, stack = F)
dev.off()
# cv > pointwise cvs 
# ==> uncertainty and sensitivity analyses will be appropriate
```



# LHS1

LHS1 is using the default settings, some interesting results were generated. Parameters seem to have been varied between 0 and 1; drawn from a uniform distribution as per the documentation. Can apply same visualisation as above to examine them. 

```{r}
#LHS1 <- LHS(rev_mod, N = 50, repetitions = 10, factors = factors, nboot = 50)
#save(LHS1, file = "attempt2/LHS1.RData")
load("LHS_proper/LHS1.RData")
```

```{r}
# prcc
png("LHS_proper/LHS1_prcc.png", res = 144, width = 1200, height = 900)
plotprcc(LHS1, ylab = res.names)
dev.off()
```

```{r}
# get.data(LHS1)
# get.N(LHS1)
# get.repetitions(LHS1)
# get.ninputs(LHS1)
# get.noutputs(LHS1)
# get.results(LHS1)
```


