---
output:
  html_document: default
---

```{r setup, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE,
                      warning = FALSE, out.extra = "")
library("GillespieSSA")
library("ggplot2")
library("patchwork")
library("ggpmisc")
library("stringr")
library("pse")
library("purrr")
library("cowplot")
library("rbenchmark")
source(file = "../dependencies/plt_fnc.R") # plotting function
```

This file contains an example of implementing a Latin-hypercube sampling based sensitivity analysis (SA) using the `pse` package framework. It starts out by spelling the reversible $CO_2$ fixation SSA model; generates a small LH sample to demonstrate the available functions. Then a more proper SA analysis is conducted (written up in the BIOL0044 project). Finally, a LH SA using the default settings is saved, as it generated interesting results. 


# Reversible CO2 fixation with equal forward and backward rates 

Writing the model out helps appreciate I/O relationships between `ssa()` and `LHS()`. 

```{r}
# initial concentrations of reactants and products involved 
nmol <- 5000

x_rev_c_fix <- c(
    CO2 = nmol,
    H2 = nmol,
    formic_acid = 0,
    formaldehyde = 0,
    methanol = 0,
    CO = 0,
    acetic_acid = 0,
    H2O = 0) 

a_rev_c_fix <- c( # reversible carbon fixation
                  "k_1*CO2*H2",
                  "k_1a*formic_acid",
                  "k_2*formic_acid*H2",
                  "k_2a*formaldehyde*H2O",
                  "k_3*formaldehyde*H2",
                  "k_3a*methanol",
                  "k_4*methanol*CO",
                  "k_4a*acetic_acid",
                  "k_5*CO2*H2",
                  "k_5a*CO*H2O")

# state change matrix
mat_rev_c_fix <- as.matrix(read.csv("../rev_c_fix/data/reversible_c_fix.csv", header = F)[-1])
```

```{r}
# parameters
parms_k <- c(k_1 = 10, k_1a = 10, 
             k_2 = 19.9, k_2a = 19.9, 
             k_3 = 23.5, k_3a = 23.5, 
             k_4 = 0.869, k_4a = 0.869, 
             k_5 = 0.061, k_5a = 0.061)

#output
out_rev_c_fix <- ssa(x0 = x_rev_c_fix, 
                      a = a_rev_c_fix, 
                      nu = mat_rev_c_fix,
                      parms = parms_k, 
                      tf = 0.0005, verbose = F, ignoreNegativeState = F)
#cat("Output dim =", dim(out_rev_c_fix$data))

rev_negl = plt_fnc(
  to_plot = colnames(out_rev_c_fix[["data"]])[-1], 
  data = out_rev_c_fix[["data"]], col = "darkred"
  )
#for (i in 1:16) cat("p1[[", i, "]] + ", sep = "")
fig = rev_negl[[1]] + rev_negl[[2]] + rev_negl[[3]] + rev_negl[[4]] + rev_negl[[5]] + rev_negl[[6]] + rev_negl[[7]] + rev_negl[[8]]
```

```{r}
# visualise results to check that sim is working 
fig
```



# PSE example with SSA model 

```{r}
# input parameters 
# params to be explored
factors = names(parms_k) # given as a vector
# their probability density distributions 
q = rep("qunif", 10) # given as a list 
# the arguments to be supplied to the probability density distributions
# given as a list of lists
q.arg = replicate(10, list(min = 1e-1, max = 1e2), simplify = F)
# names of results, to be used for plotting 
res.names = colnames(out_rev_c_fix$data)[-1]


# model 
# this function defines one iteration/run of the model 
# it takes SEPARTE vectors as arguments
# this is because the LH output is a matrix with cols = params and rows = vals 
### the arguments of this function would be modified to customise param exploration
rev_mod_run = function(k_1, k_1a, k_2, k_2a, k_3, 
                       k_3a, k_4, k_4a, k_5, k_5a){
  prms = c(k_1, k_1a, k_2, k_2a, k_3, k_3a, k_4, k_4a, k_5, k_5a)
  # since the GillespieR package takes in a named vector, make this here
  names(prms) = names(parms_k)
  # make sure the names of the parameters match 
  # with the input of the model run at the top of the document!
  run = ssa(
      x0 = x_rev_c_fix,
      a = a_rev_c_fix,
      nu = mat_rev_c_fix,
      parms = prms, # parms_k 
      # GillespieSSA package takes in a named vector of parameters 
      tf = 0.0005,
      ignoreNegativeState = F
    )
  # return the final state of the system (minus time data)
  return(as.array(run[["data"]][dim(run$data)[1],-1]))
  # can otherwise get individual columns out or some combination 
  ### editing what is returned could help tailor the sensitivity 
  ### analysis to a particular question of interest 
}


# wrapper function to run model 
# the wrapper should run the model given a data frame 
# with cols = params and rows = instances/values 
rev_mod = function(my_d) {
  names(my_d) = names(parms_k) 
  # NB name parmeters correctly, otherwise the model [ssa()] will not run 
  return(mapply(rev_mod_run, my_d[,1], my_d[,2], my_d[,3], 
                my_d[,4], my_d[,5], my_d[,6], 
                my_d[,7], my_d[,8], my_d[,9], my_d[,10]))
  }

# latin hypercube generation 
myLHS <- LHS(
  model = rev_mod, # model
  factors = factors, # parameters explored
  N = 50, # size of hypercube = number of samples generated 
  q = q, # quantile functions 
  q.arg = q.arg, # arguments for quantile functions
  nboot = 10, # bootstrap replicates; used for PRCC calculations
  res.names = res.names) # names used for visualisations
  #, repetitions = 5) # number of repetitions
```

```{r}
# checking that output and res.named match 
# as.array(out_rev_c_fix$data[dim(out_rev_c_fix$data)[1],-1])
# res.names
```

```{r}
### get data out for eyeballing 
# get.data(myLHS)
# get.repetitions(myLHS)
# get.results(myLHS)
```


# Visualise example

## ECDF 

```{r}
# examine the distribution of the model results
png("attempt2/ecdf1.png")
plotecdf(myLHS, stack = T)#, xlim=c(0,100))
  # stack = F gives several panels
    # applicable to multiple response vars
dev.off()
```

## Scatterplot

```{r}
# scatterplot of the result as a function of each parameter
pdf("attempt2/scatter.pdf")
plotscatter(myLHS)
  # index.res = which columns (aka factors) to be plotted
dev.off()
```

## PRCC 

PRCC is not very appropriate for stochastic models; PIC would be more appropriate. That being said, PIC is not easily visualisable the way PRCC is... 

```{r}
# partial rank correlation coefficient 
# assess the strength of the linear associations 
# between the result and each input parameter 
# after removing the linear effects of the other parameters
### used for deterministic models! 
purrr::map(1:7, function(x) {
  png(paste0("attempt1/prcc", x, ".png"))
  plotprcc(myLHS, index.res = x)
  dev.off()})
```

## PIC

Not easy to visualise. 

```{r}
# partial inclination coefficient
# stochastic equivalent of PRCC
# assess the sensitivity of the results to each parameter 

pic_an = pic(myLHS, nboot = 40)
# plt = myLHS
# plt$prcc = plts
# plotprcc(plt)
```

```{r}
colnames(myLHS$prcc[[1]]$PRCC)
plts = lapply(pic_an, function(x) x$prcc = x$pic)
```




# Proper PSE - included in project writeup

In the following sensitivity analysis, a Latin hypercube was used to vary parameters and examine the model outputs. 

All the parameters were set to be uniformly distributed, with range $[0.01,100]$. 

```{r}
load("LHS_proper/LHS_proper.RData")
```

## Check LH

This code helps check if the Latin hypercube samples behave as expected. 

```{r}
factors = names(parms_k) # given as a vector
q = rep("qunif", 10) # given as a list 
q.arg = replicate(10, list(min = 1e-2, max = 100), simplify = F)
res.names = colnames(out_rev_c_fix$data)[-1]
```


## LHS proper SA

```{r}
# benchmark PSE process time 
start_time <- Sys.time()
LHS_proper = LHS(model = rev_mod, factors = factors, 
                 N = 50, q = q, q.arg = q.arg, 
                 res.names = res.names, nboot = 50, repetitions = 10)
end_time <- Sys.time()
end_time - start_time
# Time difference of 14.81232 mins
```



## Plot parameters generated

```{r}
ggplot(reshape2::melt(LHS_proper$data), aes(value, 
                                       fill = variable)) + 
  geom_bar() + 
  facet_wrap(~variable)
```

```{r}
# check data
#get.data(LHS_proper)

# save data 
# save(LHS_proper, file = "attempt2/LHS_proper.RData")
```


# Plots 

```{r}
# prcc 
png("LHS_proper/proper_prcc.png", res = 144, width = 1200, height = 800)
plotprcc(LHS_proper, ylab = res.names)
dev.off()
```


```{r}
# scatterplot of the result as a function of each parameter
p = 0
for (i in seq(1,10,by = 2)) {
  p = p + 1
  png(paste0("LHS_proper/scatter",p,".png"),
  plotscatter(LHS_proper, index.data = c(i, i + 1)))
  dev.off()
}
```

```{r}
# scatterplot of the result as a function of each parameter
plotss_scatter = map(seq(1,10,by = 2), function(x)
  return(eval(parse(
    text = paste0(
      "~{plotscatter(LHS_proper, index.data = c(", 
      paste(x), 
      ", ", paste(x + 1),"))}"
    )))))
```

```{r}
# scatter plots
png("LHS_proper/proper_scatters.png", res = 100, width = 500, height = 2200)
plot_grid(
  ggdraw(plotss_scatter[[1]]),
  ggdraw(plotss_scatter[[2]]),
  ggdraw(plotss_scatter[[3]]),
  ggdraw(plotss_scatter[[4]]),
  ggdraw(plotss_scatter[[5]]),
  nrow = 5, ncol = 1)
dev.off()
```


```{r}
# examine the distribution of the model results
plotss = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotecdf(LHS_proper, stack = F, xlab = res.names[",
      paste(x),
      "], index.res = ",
      paste(x),
      ")}"
    )
  ))))

# for (i in 1:8) {
#   cat(paste0("ggdraw(plotss[[",i,"]]), "))
#   ggdraw(plotss[[i]])
# }
```

```{r}
# ecdfs 
png("LHS_proper/proper_ecdfs.png", res = 100, width = 1000, height = 600)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),
  ggdraw(plotss[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]), 
  nrow = 2, ncol = 4)
dev.off()
```


```{r}
# example of getting globsl cv (coefficient of)
r <- get.results(LHS_proper)
sd(r) / mean(r) # global CV
plotcv(LHS_proper, stack = F, index.res = 1)
```

```{r}
plotss_cv = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotcv(LHS_proper, stack = F, index.res = ", paste(x),")}"
    )
  ))))
```

```{r}
library("cowplot")
png("LHS_proper/proper_cv.png", res = 100, width = 1000, height = 600)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),
  ggdraw(plotss[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]), 
  nrow = 2, ncol = 4, labels = res.names, vjust = 3.5)
dev.off()
```


```{r}
# mixed plot with ecdfs and cvs one below the other 
plotss = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotecdf(LHS_proper, stack = F, xlab = \"ecdf\", index.res = ",
      paste(x), ")}"
    )
  ))))

plotss_cv = map(1:8, function(x)
  return(eval(parse(
    text = paste0(
      "~{plotcv(LHS_proper, stack = F, index.res = ", paste(x),")}"
    )
  ))))

res.names.new = c(rep(res.names[1:4], 2), rep(res.names[5:8], 2))

png("LHS_proper/proper_ecdf_cv.png", res = 100, width = 1000, height = 1200)
plot_grid(
  ggdraw(plotss[[1]]),
  ggdraw(plotss[[2]]),
  ggdraw(plotss[[3]]),  
  ggdraw(plotss[[4]]),
  ggdraw(plotss_cv[[1]]),
  ggdraw(plotss_cv[[2]]),
  ggdraw(plotss_cv[[3]]),
  ggdraw(plotss_cv[[4]]),
  ggdraw(plotss[[5]]),
  ggdraw(plotss[[6]]),
  ggdraw(plotss[[7]]),
  ggdraw(plotss[[8]]),
  ggdraw(plotss_cv[[5]]),
  ggdraw(plotss_cv[[6]]),
  ggdraw(plotss_cv[[7]]),
  ggdraw(plotss_cv[[8]]),
  nrow = 4, ncol = 4, labels = res.names.new, vjust = 3.5)
dev.off()
```

```{r}
# cv 
png("LHS_proper/proper_cv.png", res = 144, width = 1000, height = 600)
plotcv(LHS_proper, stack = F)
dev.off()
# cv > pointwise cvs 
# ==> uncertainty and sensitivity analyses will be appropriate
```



# LHS1

LHS1 is using the default settings, some interesting results were generated. Parameters seem to have been varied between 0 and 1; drawn from a uniform distribution as per the documentation. Can apply same visualisation as above to examine them. 

```{r}
#LHS1 <- LHS(rev_mod, N = 50, repetitions = 10, factors = factors, nboot = 50)
#save(LHS1, file = "attempt2/LHS1.RData")
load("LHS_proper/LHS1.RData")
```

```{r}
# prcc
png("LHS_proper/LHS1_prcc.png", res = 144, width = 1200, height = 900)
plotprcc(LHS1, ylab = res.names)
dev.off()
```

```{r}
# get.data(LHS1)
# get.N(LHS1)
# get.repetitions(LHS1)
# get.ninputs(LHS1)
# get.noutputs(LHS1)
# get.results(LHS1)
```


